---
title: "PA1_template"
author: "Hemant"
output: html_document
---


```{r, echo=FALSE}
require(ggplot2)
require(plyr)
require(dplyr)
```

## Importing the file activity.csv

```{r setup, echo=TRUE}
activity<-read.csv("./repdata_data_activity/activity.csv", sep=",", header = TRUE)
activity$date<-as.Date(activity$date)
head(activity)
```

## Plotting histogram of total steps per day

```{r histogram}
#library(ggplot2)
agg<-aggregate(steps~date, activity, sum)
g<-ggplot(agg, aes(steps))
g+geom_histogram(color="black", fill="#03FC49")
```

## Mean and Median of the steps taken each day.

```{r}
mean(activity$steps, na.rm=TRUE)
median(activity$steps, na.rm=TRUE)
```

## Mean and Median of the total steps taken each day.

```{r}
mean(agg$steps, na.rm=TRUE)
median(agg$steps, na.rm=TRUE)
```

## Time series plot of Steps vs Intervals 

```{r}
agg<-aggregate(steps~interval, activity, mean, na.rm=TRUE)
g<-ggplot(agg, aes(interval, steps))
g+geom_point(color="#0398FC")+geom_line(color="#FC0328")
```

## Interval with the highest number of steps

```{r}
agg[agg$steps==max(agg$steps),]$interval
```

## Imputing NA values in the dataset with the mean values of steps taken in a new dataset

```{r imputing}
activity_na<-activity
activity_na[is.na(activity_na)]=mean(activity_na$steps, na.rm=TRUE)
head(activity_na)
```

## Histogram of total steps taken each day

```{r}
agg<-aggregate(steps~date, activity_na, sum)
g<-ggplot(agg, aes(steps))
g+geom_histogram(color="black", fill="#0390FC")
```

## Mean and Median after imputing NA values

```{r}
mean(agg$steps, na.rm=TRUE)
median(agg$steps, na.rm=FALSE)
```

The aggregated mean is same as before but the median differs. This is due to the presence of 37.38 in place of NA values in the dataset.

## Using weekdays() analysis comparing average steps taken per 5-minute interval on weekends and weekdays 

```{r}
activity_na<-mutate(activity_na, day=weekdays(date))
head(activity_na)
sub_weekday<-filter(activity_na, day!=c("Saturday","Sunday"))
sub_weekend<-filter(activity_na, day==c("Saturday","Sunday"))
sub_weekday$day<-"Weekday"
sub_weekend$day<-"Weekend"
activity_day<-rbind(sub_weekday, sub_weekend)
agg<-aggregate(steps~interval+day, activity_day, mean)
g<-ggplot(agg, aes(interval, steps))
g+geom_point(color="#75192E")+geom_line(color="#1057C2")+facet_grid(day~.)
```
